True versus apparent shapes of bow shocks and bow waves

### 2017-11-12 papers/Paper1/app-shape-parameters.tex
Finish sec 5.2 on true shapes of cantoids and ancantoids
Re-do the projected ancantoid fig and start describing it

To find the alatude, \(\Lambda = R_{90} / R_0\), we use equation~\eqref{eq:crw-angles} at \(\theta = 90^\circ\) to write
\begin{equation}
  \label{eq:Lambda-from-theta-1-90}
  \Lambda = \frac{D} {R_0} \tan \theta_{1,90} \ , 
\end{equation}
where \(\theta_{1,90} = \theta_1(\theta = 90^\circ)\), which, following
equation~\eqref{eq:ancantoid-theta-theta1-implicit}, must satisfy
\begin{equation}
  \label{eq:theta-1-90-implicit}
  \theta_{1,90} \cot \theta_{1,90}  = 1 - \frac{2 \beta}{k + 2} \ . 
\end{equation}
Combining \eqref{eq:Lambda-from-theta-1-90} and
\eqref{eq:theta-1-90-implicit} with \eqref{eq:again-R0-over-D} yields
\begin{equation}
  \label{eq:Lambda-beta-xi-theta-1-90}
  \Lambda = \frac{ \left(1 + \beta^{1/2}\right) \,\theta_{1,90}} {\beta^{1/2} \left(1 - \xi_k \beta\right)} \ ,
\end{equation}
where
\begin{equation}
  \label{eq:xi-k}
  \xi_k = \frac{2} {k +2} \ .
\end{equation}
We now take the Taylor expansion of equation~\eqref{eq:theta-1-90-implicit} to find
\begin{equation}
  \label{eq:theta-1-90-Taylor}
  \theta_{1,90}^2 + \tfrac{1}{15}  \theta_{1,90}^4 + \mathcal{O}(\theta_{1,90}^6)
  = 3 \xi_k \beta \ , 
\end{equation}
which, if \(\theta_{1,90}\) is small, has the approximate solution
\begin{equation}
  \label{eq:theta-1-90-approx}
  \theta_{1,90} \approx \left( \frac{3 \xi_k \beta} {1 + \tfrac15 \xi_k \beta} \right) \ .
\end{equation}
Substituting back into equation~\eqref{eq:Lambda-from-theta-1-90}
yields an approximate value for the alatude of
\begin{equation}
  \label{eq:Lambda-approx}
  \text{ancantoid} \quad
  \Lambda \approx \frac {(3 \xi_k)^{1/2} \left( 1 + \beta^{1/2} \right)}
  { \left( 1 + \tfrac15 \xi_k \beta \right)^{1/2} \left( 1 - \xi_k \beta \right)} \ .
\end{equation}
This approximation is surprisingly accurate, with a relative error of
order \(1\%\) even for \(\beta\) as large as \(0.5\) with \(k = 0\).  The
leading term in the relative error scales as \((\beta / (k + 2))^2\), so the
approximation is even better for smaller \(\beta\) and larger \(k\).

@@

The asymptotic opening angle
of the far wings, \(\theta_\infty\), can be found from
equation~\eqref{eq:ancantoid-theta-theta1-implicit} for the
ancantoids, together with the condition that
\(\theta_\infty + \theta_{1\infty} = \pi\).  These yield the implicit equation
\begin{equation}
  \label{eq:anacantoid-theta-inf}
  \theta_\infty - \left( \frac {k + 2 (1 - \beta)} {k + 2} \right) \tan \theta_\infty
  = \pi + 2 \beta I_k(\pi/2) \ ,
\end{equation}
where
\begin{equation}
  I_k(\pi/2) = \frac{\sqrt \pi}{4} \,
      \frac{  \GammaFunc\left( \frac{k+1}{2} \right)} {\GammaFunc\left(\frac{k+4}{2}\right)}
\end{equation}
and \(\GammaFunc\) is the usual Gamma function.  This can be compared with the equivalent result obtained by \CRW{} for the cantoids:
\begin{equation}
  \label{eq:cantoid-theta-inf}
  \theta_\infty - \tan \theta_\infty = \frac{\pi}{1 - \beta} \ .
\end{equation}
Note that, unlike in the cases of \(\Pi\) and \(\Lambda\),
equation~\eqref{eq:anacantoid-theta-inf} does \emph{not} reduce to
equation~\eqref{eq:cantoid-theta-inf} in the limit \(k \to 0\).  This is
because, for \(\theta > 90^\circ\), the \(k = 0\) ancantoid differs from the
cantoid since the former has no wind in the backward hemisphere (see
Figure~\ref{fig:anisotropic-arrows}).  Therefore there is less inner
support for the far wings of the bow, and so \(\theta_\infty\) is smaller than
in the cantoid case.  The wilkinoid result again follows from
\(\beta \to 0\), implying that \(\theta_\infty = \pi\), or, in other words, that the far
wings are asymptotically parallel to the symmetry axis, as is the case
for the paraboloid (App.~\ref{app:parabola}).  In the case of the
wilkinoid, however, the behavior is cubic in the wings,
\(z \sim r^3\), as opposed to quadratic as in the paraboloid.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "quadrics-bowshock"
%%% End:

>>>>

To find the alatude, \(\Lambda = R_{90} / R_0\), we use equation~\eqref{eq:crw-angles} at \(\theta = 90^\circ\) to write
\begin{equation}
  \label{eq:Lambda-from-theta-1-90}
  \Lambda = \frac{D} {R_0} \tan \theta_{1,90} \ , 
\end{equation}
where \(\theta_{1,90} = \theta_1(\theta = 90^\circ)\), which, following
equation~\eqref{eq:ancantoid-theta-theta1-implicit}, must satisfy
\begin{equation}
  \label{eq:theta-1-90-implicit}
  \theta_{1,90} \cot \theta_{1,90}  = 1 - \frac{2 \beta}{k + 2} \ . 
\end{equation}
Combining \eqref{eq:Lambda-from-theta-1-90} and
\eqref{eq:theta-1-90-implicit} with \eqref{eq:again-R0-over-D} yields
\begin{equation}
  \label{eq:Lambda-beta-xi-theta-1-90}
  \Lambda = \frac{ \left(1 + \beta^{1/2}\right) \,\theta_{1,90}} {\beta^{1/2} \left(1 - \xi_k \beta\right)} \ ,
\end{equation}
where
\begin{equation}
  \label{eq:xi-k}
  \xi_k = \frac{2} {k +2} \ .
\end{equation}
We now take the Taylor expansion of equation~\eqref{eq:theta-1-90-implicit} to find
\begin{equation}
  \label{eq:theta-1-90-Taylor}
  \theta_{1,90}^2 + \tfrac{1}{15}  \theta_{1,90}^4 + \mathcal{O}(\theta_{1,90}^6)
  = 3 \xi_k \beta \ , 
\end{equation}
which, if \(\theta_{1,90}\) is small, has the approximate solution
\begin{equation}
  \label{eq:theta-1-90-approx}
  \theta_{1,90} \approx \left( \frac{3 \xi_k \beta} {1 + \tfrac15 \xi_k \beta} \right) \ .
\end{equation}
Substituting back into equation~\eqref{eq:Lambda-from-theta-1-90}
yields an approximate value for the alatude of
\begin{equation}
  \label{eq:Lambda-approx}
  \text{ancantoid} \quad
  \Lambda \approx \frac {(3 \xi_k)^{1/2} \left( 1 + \beta^{1/2} \right)}
  { \left( 1 + \tfrac15 \xi_k \beta \right)^{1/2} \left( 1 - \xi_k \beta \right)} \ .
\end{equation}
This approximation is surprisingly accurate, with a relative error of
order \(1\%\) even for \(\beta\) as large as \(0.5\) with \(k = 0\).

@@

The asymptotic opening angle
of the far wings, \(\theta_\infty\), can be found from
equation~\eqref{eq:ancantoid-theta-theta1-implicit} for the
ancantoids, together with the condition that
\(\theta_\infty + \theta_{1\infty} = \pi\).  These yield the implicit equation
\begin{equation}
  \label{eq:ancantoid-theta-inf}
  \theta_\infty - \left( \frac {k + 2 (1 - \beta)} {k + 2} \right) \tan \theta_\infty
  = \pi + 2 \beta I_k(\pi/2) \ ,
\end{equation}
where
\begin{equation}
  I_k(\pi/2) = \frac{\sqrt \pi}{4} \,
      \frac{  \GammaFunc\left( \frac{k+1}{2} \right)} {\GammaFunc\left(\frac{k+4}{2}\right)}
\end{equation}
and \(\GammaFunc\) is the usual Gamma function.  This can be compared with the equivalent result obtained by \CRW{} for the cantoids:
\begin{equation}
  \label{eq:cantoid-theta-inf}
  \theta_\infty - \tan \theta_\infty = \frac{\pi}{1 - \beta} \ .
\end{equation}
Note that, unlike in the cases of \(\Pi\) and \(\Lambda\),
equation~\eqref{eq:ancantoid-theta-inf} does \emph{not} reduce to
equation~\eqref{eq:cantoid-theta-inf} in the limit \(k \to 0\).  This is
because, for \(\theta > 90^\circ\), the \(k = 0\) ancantoid differs from the
cantoid since the former has no wind in the backward hemisphere (see
Figure~\ref{fig:anisotropic-arrows}).  Therefore there is less inner
support for the far wings of the bow, and so \(\theta_\infty\) is smaller than
in the cantoid case.  The wilkinoid result again follows from
\(\beta \to 0\), implying that \(\theta_\infty = \pi\), or, in other words, that the far
wings are asymptotically parallel to the symmetry axis, as is the case
for the paraboloid (App.~\ref{app:parabola}).  In the case of the
wilkinoid, however, the behavior is cubic in the wings,
\(z \sim r^3\), as opposed to quadratic as in the paraboloid.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "quadrics-bowshock"
%%% End:

### 2017-11-12 papers/Paper1/sec-thin-shell.tex
Re-do the projected ancantoid fig and start describing it

%The most notable scenarios, since they have astrophysical relevance are the following:
%$k=1/2$ ak.a. the ``proplyd case'', following \citep{HA:1998}, and $k=0$, ak.a, the ``isotropic case'', following \CRW{}. The comparison between both solutions
%is shown in figure (\ref{fig:r-beta}), along with an extreme anisotropy case.

@@

However, the true planitude and alatude, which are what would be
observed for a side-on viewing angle (\(i = 0\)), are not at all
sensitive to the behavior of the far wings of the bow shock, which has
a rather different implication as to which variety of quadric best
approximates each shape.  We illustrate this is
Figure~\ref{fig:ancantoid-angles}, which shows two different ways of
estimating the quadric angle, \(\theta_{\Q}\) (see \S~\ref{sec:conic}).
The first is from \((\Pi, \Lambda)\), as in
Figure~\ref{fig:ancantoid-Pi-lambda-true}:
\newcommand\head{^{\text{head}}}
\newcommand\tail{^{\text{tail}}}
\begin{equation}
  \label{eq:thetaQ-head}
  \theta_{\Q}\head =
  \sgn{\bigl(2 \Pi - \Lambda^2\bigr)} \;
  \tan^{-1} \Abs{2\Pi - \Lambda^2}^{1/2} \ ,
\end{equation}
which follows from equations~\eqref{eq:Tq}, \eqref{eq:thetaQ}, and
\eqref{eq:Tq-from-Pi-Lambda}.  The second is from the asymptotic
opening angle of the wings, \(\theta_\infty\) (Fig.~\ref{fig:characteristic-radii}):
\begin{equation}
  \label{eq:thetaQ-tail}
  \theta_{\Q}\tail = \theta_\infty - 180^\circ \ , 
\end{equation}
where \(\theta_\infty\) is calculated from
equation~\eqref{eq:anacantoid-theta-inf} for ancantoids, or
\eqref{eq:cantoid-theta-inf} for cantoids.  If the bow shock shape
were truly a quadric, then these two definitions would agree.
However, as seen in Figure~\ref{fig:ancantoid-angles}, this is not the
case for the cantoids and ancantoids.  While
\(\smash[b]{\theta_{\Q}\head}\) generally corresponds to a prolate spheroid
(except for the largest values of \(\beta\)),
\(\smash[b]{\theta_{\Q}\tail}\) always corresponds to a hyperbola.  This
has important implications for the projected shapes (as we will see in
the next section), since the far wings influence the projected
planitude and alatude when the inclination is large.

@@

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.8\linewidth]{figs/test_xyprime}
  \caption{Apparent bow shapes as a function of inclination angle for
    isotropic thin shell models. (a)~Confocal paraboloid for
    comparison (shape independent of inclination).
    (b)~Wilkinoid. (c)~Cantoid, \(\beta = 0.001\). (d)~Cantoid,
    \(\beta = 0.01\). }
  \label{fig:xyprime}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.8\linewidth]{figs/test_xyprime_ancantoid}
  \caption{Further apparent bow shapes as a function of inclination
    angle for anisotropic thin shell models (ancantoids).
    (a)~\(\beta = 0.001\), \(\xi =.8\); (b) (a)~\(\beta = 0.1\),
    \(\xi =.8\); (c)~\(\beta = 0.001\) \(\xi =0.4\); (d)
    \(\beta = 0.1\) \(\xi =0.4\).}
  \label{fig:xyprime-anacantoid}
\end{figure*}

\begin{figure*}
  \centering
  \setkeys{Gin}{width=0.48\linewidth}
  \begin{tabular}{@{}ll@{}}
    (a) & (b) \\
    \includegraphics{figs/ancantoid-R90-vs-Rc-a}
        & \includegraphics{figs/ancantoid-R90-vs-Rc-b} \\
  \end{tabular}
  \caption{Diagnostic diagram for thin shell models}
  \label{fig:thin-shell-R90-Rc}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.8\textwidth]{figs/two-quadric-R90-Rc-snapshots}
  \caption{Variation with inclination angle of radii diagnostics}
  \label{fig:thin-shell-R90-Rc-snapshots}
\end{figure*}

>>>>

\begin{figure*}
  \centering
  \includegraphics[width=0.9\linewidth]{figs/test_xyprime}
  \caption{Apparent bow shapes as a function of inclination angle for
    isotropic thin shell models. (a)~Confocal paraboloid for
    comparison (shape independent of inclination).
    (b)~Wilkinoid. (c)~Cantoid, \(\beta = 0.001\). (d)~Cantoid,
    \(\beta = 0.01\). }
  \label{fig:xyprime}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.9\linewidth]{figs/test_xyprime_ancantoid}
  \caption{Further apparent bow shapes as a function of inclination
    angle for anisotropic thin shell models (ancantoids).
    (a)~\(\beta = 0.001\), \(k = 0.5\); (b) (a)~\(\beta = 0.1\),
    \(k = 0.5\); (c)~\(\beta = 0.001\) \(k = 3\); (d)
    \(\beta = 0.1\) \(k = 3\).}
  \label{fig:xyprime-ancantoid}
\end{figure*}

@@

However, the true planitude and alatude, which are what would be
observed for a side-on viewing angle (\(i = 0\)), are not at all
sensitive to the behavior of the far wings of the bow shock, which has
a rather different implication as to which variety of quadric best
approximates each shape.  We illustrate this is
Figure~\ref{fig:ancantoid-angles}, which shows two different ways of
estimating the quadric angle, \(\theta_{\Q}\) (see \S~\ref{sec:conic}).
The first is from \((\Pi, \Lambda)\), as in
Figure~\ref{fig:ancantoid-Pi-lambda-true}:
\newcommand\head{^{\text{head}}}
\newcommand\tail{^{\text{tail}}}
\begin{equation}
  \label{eq:thetaQ-head}
  \theta_{\Q}\head =
  \sgn{\bigl(2 \Pi - \Lambda^2\bigr)} \;
  \tan^{-1} \Abs{2\Pi - \Lambda^2}^{1/2} \ ,
\end{equation}
which follows from equations~\eqref{eq:Tq}, \eqref{eq:thetaQ}, and
\eqref{eq:Tq-from-Pi-Lambda}.  The second is from the asymptotic
opening angle of the wings, \(\theta_\infty\) (Fig.~\ref{fig:characteristic-radii}):
\begin{equation}
  \label{eq:thetaQ-tail}
  \theta_{\Q}\tail = \theta_\infty - 180^\circ \ , 
\end{equation}
where \(\theta_\infty\) is calculated from
equation~\eqref{eq:ancantoid-theta-inf} for ancantoids, or
\eqref{eq:cantoid-theta-inf} for cantoids.  If the bow shock shape
were truly a quadric, then these two definitions would agree.
However, as seen in Figure~\ref{fig:ancantoid-angles}, this is not the
case for the cantoids and ancantoids.  While
\(\smash[b]{\theta_{\Q}\head}\) generally corresponds to a prolate spheroid
(except for the largest values of \(\beta\)),
\(\smash[b]{\theta_{\Q}\tail}\) always corresponds to a hyperbola.  This
tension between the shape of the head and the shape of the far wings
has important implications for the projected shapes (as we will see in
the next section), since the far wings influence the projected
planitude and alatude when the inclination is large.

@@

\begin{figure}
  \centering
  \setkeys{Gin}{width=\linewidth}
  \begin{tabular}{@{}l@{}}
    (a) \\
    \includegraphics{figs/ancantoid-R90-vs-Rc-a} \\
    (b) \\
    \includegraphics{figs/ancantoid-R90-vs-Rc-b}
  \end{tabular}
  \caption{Apparent projected shapes of wilkinoid, cantoids and
    ancantoids in the \(\Pi'\)--\(\Lambda'\) plane.  Colored dots indicate the
    \(\abs{i} = 0\) position (as shown in
    Fig.~\ref{fig:ancantoid-Pi-lambda-true}).}
  \label{fig:thin-shell-R90-Rc}
\end{figure}

@@

Figures~\ref{fig:xyprime} and~\ref{fig:xyprime-ancantoid} show the
apparent bow shapes of various thin shell models (wilkinoid, cantoids,
ancantoids) for different inclination angles \(\abs{i}\).  For
comparison, Figure~\ref{fig:xyprime}a shows the confocal paraboloid,
whose apparent shape is independent of inclination (see
Appendix~\ref{app:parabola}).  The wilkinoid (Fig.~\ref{fig:xyprime}b)
shows only subtle changes, with the wings becoming slightly more swept
back as the inclination increases.  The cantoids
(Fig.~\ref{fig:xyprime}c and d) behave in the opposite way, with the
wings becoming markedly more open once \(\abs{i}\) exceeds
\(60^\circ\) (for \(\beta = 0.001\)), or \(45^\circ\) (for
\(\beta = 0.01\)).  The ancantoids (Fig.~\ref{fig:xyprime-ancantoid}) can
show more complex behavior.  For instance, in the \(k = 0.5\),
\(\beta = 0.001\) ancantoid (Fig.~\ref{fig:xyprime-ancantoid}a) the near
wings begin to become more closed with increasing inclination up to
\(\abs{i} = 60^\circ\), at which point they open up again.

% \begin{figure*}
%   \centering
%   \includegraphics[width=0.8\textwidth]{figs/two-quadric-R90-Rc-snapshots}
%   \caption{Variation with inclination angle of radii diagnostics}
%   \label{fig:thin-shell-R90-Rc-snapshots}
% \end{figure*}
